---
title: "WSDC Project (part 1)"
author: "Connie Wang"
date: "December 28, 2017"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpmisc)
library(cowplot)
knitr::opts_chunk$set(fig.path='Figs/')
```

### **Revisions: 1/4/2018**
Hi! After I posted, two things were pointed out to me that changed the results a little:

1. Many people start in Newcomer. I changed my filter from collecting only those with at least one Novice point to collecting those with at least one Novice or Newcomer point. I also changed my time-by-region graphs to look at date of first record (whether Newcomer or Novice, where before the first record was necessarily Novice)
2. The rule change that modified required Novice points from 20 to 15 happened at the start of 2012. I changed the filter in the section that looks at Novices to include data from 2012 onwards.

I applied the changes throughout the R and Tableau; here's a summary of what's different:

1. 9% of current Intermediate (not 12%) have exactly 15 Novice points and no Intermediate points
2. 38% of current Intermediate points (not 50%) have between 15 and 30 Novice points and no Intermediate points
3. When comparing those with 15 Novice points and those with more than 15 Novice points, the difference in how often they place 1st-5th at a tier 3 or 1st-3rd at a tier 2 in Intermediate is reduced. The p-value in the other region is now 0.068. All regions still trend in the same direction as they did before.

It was also pointed out to me many times that Novices with 15 points currently (that they earned prior to 2018) would be "grandfathered in" -- if you are one of these Novices and you want to compete in Intermediate, you should file a petition with the event you're competing at, and this petition will be automatically approved.

## **Introduction**
Saying that West Coast Swing is my primary hobby might be an understatement at this point -- I dance one or two nights a week, attend a convention about once a month, and practice... maybe not as often as I should, but sometimes. Many of my friends are ones I met through dance, and sometimes we talk about West Coast Swing even when we're not dancing. Suffice it to say, I end up thinking about this dance a lot (I know, I'm really cool).

In my 4 years of dancing, I've often had passing wonderings about various aspects of the dance. Luckily, the World Swing Dance Council (WSDC) tracks competition results for West Coast Swing, so there's plenty of data to play with -- conveniently, that makes good practice material for my learning and exploration in the world of data science. There are too many of these questions to cover in one post (...and I don't know how to do everything I want to do yet), but here's the first installment! It's mostly inspired by discussions around the rule changes that take effect in 2018.

As an aside: I'm pretty new to this, so I'm sure this is incredibly poorly-written code. I was pretty hesitant to make this post at all, but the topic is sort of timely, and I can't improve if I don't put myself out there, right? So if you are a computery person who is feeling generous, I would love if you could take a look and help me destupidify my code (Python and R both)! [You can find all files used (including the version of this page that doesn't hide code chunks) here](https://github.com/conniewang3/WSDC-Project/tree/master/part1). Shoot me a message -- I'd be happy to buy you lunch/coffee for your help, or skilltrade a dance lesson or something. 


## **TL;DR**
(AKA the part where I practice Tableau dashboarding -- try hovering and clicking on things!)
<iframe src="https://public.tableau.com/views/WSDCProjectPart1/TLDR?:embed=y&:display_count=yes&publish=yes&:toolbar=no&:showVizHome=no"
 width="810" height="1240" frameBorder="0"></iframe>
 
 
### **West Coast Swing**
#### (A brief primer for anyone reading this who's not already familiar with West Coast Swing)
West Coast Swing (WCS) is a versatile partner dance that emphasizes, among other things, the ability to create a dance with a partner on the spot. There is a competitive aspect of this dance which tests this ability; it involves going to weekend-long events and participating in Jack & Jills, the main competition format. In a Jack & Jill prelim round, each competitor dances 3 times, each time with a different random partner to a random song. Judges then select their favorite dancers to move on to finals, where competitors are again randomly paired and receive a placement depending on how well the judges score each partnership. Points are assigned if you final or place according to the WSDC rules, and tracked on the WSDC database. There are 4 main competition divisions: Novice, Intermediate, Advanced, and All-Star -- to move up to a higher division, a competitor must earn a certain number of points in their division. [The rules for this are outlined here.](github.com/conniewang3/WSDC-Project/blob/master/files/rules_new.pdf)


### **Motivation**
Recently, there was a change in the rules, which has sparked a lot of discussion especially among more isolated communities, like Asia and Australia. Although Europe and the US now have many WCS events per year, Asia and Australia are still smaller WCS scenes. The lack of local events and expense of traveling to the established, larger, and distant events in the US and Europe make it significantly more difficult for dancers in this scene to move up through the divisions. The rule change is perceived as making it even more difficult (or perhaps even impossible) for these dancers. Many of the questions here were inspired by the discussion around this change. [The new version of the rules can be found here.](github.com/conniewang3/WSDC-Project/blob/master/files/rules_new.pdf)


## **Results**

### **Impact of change in All-Star requirements**
Under the old rules, anyone with 45+ Advanced points within the last 3 years or 1+ All-Star points ever is eligible to compete in All-Star. Starting 2018, anyone with 45+ Advanced points or 3+ All-Star points in the last 3 years is eligible. This means that dancers with 1+ All-Star point ever but less than 3 All-Star points and less than 45 Advanced points in the last 3 years are "demoted" back to Advanced. How many active competitors will that affect?

I'll define "active" competitors in this post as having gotten a point in any division from Novice to All-Star within the last 3 years. I can think of a handful of dancers who still travel and compete in All-Star, but haven't gotten a point in several years; however, I'm guessing there are fewer of these than dancers who haven't gotten a point in 3 years because they stopped competing.


#### **How many All-Star dancers will be demoted?**

```{r new_als, warning=FALSE, echo=FALSE}
# Read in data and filter out those who are not active (no nov-als points in the last 3 years)
data <- tbl_df(read.csv('data_with_newcomer.csv', as.is=TRUE)) %>%
  filter(as.Date(paste(end_date, '01'), '%B %Y %d') > as.Date('2015-01-01'))

# Filter by division and number of All-Star points in 3 years to count those who are demoted
demoted <- nrow(data %>% filter(division == 'All-Star', als_3y < 3))

# Get count as a percentage of total All-Star and All-Star-eligible dancers
total_als <- nrow(data %>% filter(division == 'All-Star'))
als_adv <- nrow(data %>% filter(adv_3y >= 45))
total_elig <- total_als + als_adv
percents <- c(round(demoted/total_als*100, 0),
              round(demoted/total_elig*100, 0),
              round((1 - total_als/total_elig)*100, 0))
```

A total of **`r demoted`** dancers currently competiting in All-Star will need to compete in Advanced starting in 2018 (unless they successfully petition).

As All-Star is an opt-in division, I counted only those with All-Star points as "All-Star" in my dataset. The demoted dancers represent **`r percents[1]`%** of these dancers.

Those with at 45+ Advanced points in the last 3 years but 0 All-Star points can choose whether they want to compete in Advanced or All-Star -- there are **`r als_adv`** of these dancers and they constitute **`r percents[3]`%** of all those eligible to compete in All-Stars. The demoted dancers represent **`r percents[2]`%** of All-Star-*eligible* dancers.

As regional variation is kind of the theme of this post, I broke down the percentage of All-Star-eligible dancers who will be demoted by region. Note that WSDC doesn't keep data on where each dancer lives, so region data was inferred based on what events the dancer earned points at. Midwest includes Texas, Europe includes Israel, and Other is Asia/Australia/NZ/Brazil (more details at the bottom if you're interested).

```{r als_byregion, echo=FALSE}
# Find total All-Star-eligible by region
total_region <- data %>%
  filter(division == 'All-Star' | adv_3y >= 45) %>%
  count(region)

# Filter to demoted All-Stars, count and take as % of eligible in each region
als_byregion <- data %>% 
  filter(division == 'All-Star', als_3y < 3) %>% 
  count(region) %>%
  mutate(total = total_region$n) %>%
  mutate(percent = n/total*100) %>%
  mutate(SE = sqrt((n/total*(1-n/total))/total)*100)

# Plot data using ggplot2
ggplot(als_byregion, aes(x=region, y=percent)) +
  # Error bars show standard error
  geom_errorbar(aes(ymin=percent-SE, ymax=percent+SE, width = 0.1)) +
  geom_bar(aes(fill=region), position='dodge', stat='identity') +
  xlab('Region') + ylab('% All-Stars demoted') +
  ggtitle('Proportion of All-Star-eligible dancers demoted') +
  labs(fill="Region")

# Chi-sq test (is having 15 points related to region?)
# Note that several of the expected values are under 5, so I ran Fisher's exact test too (which should be robust to small size/low incidence). It takes forever and a half because we need such a large workspace, due to the sample size and p-value both being low (Sorry to anyone trying to check my work)
test_data <- als_byregion %>%
  mutate(yes = n) %>%
  mutate(no = total - n) %>%
  mutate(region = c(1,2,3,4,5)) %>%
  select(region, yes, no) %>%
  data.matrix()
fisher <- fisher.test(test_data, simulate.p.value=TRUE, B=1e7)
conf <- round(100*(1-fisher$p.value), 2)
p <- round(100*(1-conf), 2)
```
There is definitely some regional variation here -- it looks like a relatively larger portion of current All-Stars will be demoted in the East Coast and Midwest regions (error bars show standard error). 

Since repeated two-way tests would overestimate significance, I used Fisher's exact test to measure independence of the two variables instead. The p-value from this test indicates a `r conf`% confidence that the the two variables are not independent (or, to be more technically correct, that there's a 0.06% chance we'd see these results or more extreme if there were no difference in population means -- take that, pedantic naysayers). In other words, we can say with statistically-backed conviction that **the percentage of All-Star-eligible dancers who will be demoted is different in different regions.**


### **Impact of change in required number of Novice points (from 15 to 16 points)**
Under the old rules, dancers with 15+ Novice points must compete in Intermediate. With the rule change, dancers with 16+ Novice points are eligible for Intermediate, and dancers with 30+ Novice points must compete in Intermediate. To understand the impact of this rule change, I wanted to find out:

1. How many dancers will be "demoted" by the change? How many dancers will be allowed to return to Novice as a result of the change? 
2. Do dancers with exactly 15 Novice points do worse in Intermediate than those with more than 15?

To answer these questions, I looked at only dancers who pointed out of
Novice after February 2013 (when the point requirement changed from 20 to 15 points).

#### **How many dancers will be "demoted" and how many can now choose to dance down?**

```{r new_novice, echo=FALSE}
# Filter out those who pointed out of Novice before January 2012
new_novice <- data %>%
  filter(as.Date(paste(nov_end, '01'), '%B %Y %d') > as.Date('2012-02-01'))

# Filter by number of Novice and Intermediate points to count affected dancers 
must_nov <- nrow(new_novice %>% filter(nov_points == 15, int_points == 0))
can_nov <- nrow(new_novice %>% filter(nov_points %in% seq(15,29), int_points == 0))

# Get counts as a percentage of total Novice and total Intermediate dancers
total_nov <- nrow(new_novice %>% filter(division == 'Novice'))
total_int <- nrow(new_novice %>% filter(division == 'Intermediate'))
percents <- c(round(must_nov/total_nov*100, 0),
              round(can_nov/total_nov*100, 0),
              round(must_nov/total_int*100, 0),
              round(can_nov/total_int*100, 0))
```

Of people who were previously competing in Intermediate, **`r must_nov`** now *must* compete in Novice (unless they submit a petition, which will be automatically approved), and **`r can_nov`** now *have the option* to compete in Novice. This represents **`r percents[1]`%** and **`r percents[2]`%** of current Novice dancers or **`r percents[3]`%** and **`r percents[4]`%** of current Intermediate dancers. 

I'm pretty surprised by how high these numbers are, actually. I wonder how many of these dancers are actually in the presumed target group of dancers who compete only once in Novice... (quiet sounds of number-crunching in the background)

```{r just_wondering, echo=FALSE}
one_comp <- nrow(filter(new_novice, nov_count == 1, nov_points == 15, int_points == 0))
percent <- round(one_comp/must_nov*100, 0)

# I think this is binomially distributed, so it should be appropriate to use the mean
mean_comps <- round(mean(filter(new_novice, nov_points == 15, int_points == 0)$nov_count), 0)
```

Wow, I totally overestimated the number of Novice dancers who got their 15 points exactly all in one go. I really thought the change from 15 to 16 points was targeted at dancers who get one lucky draw at a tier 3 comp, place 1st, and are immediately out. Turns out, that's only **`r percent`%** of total affected dancers (those with exactly 15 Novice points and 0 Intermediate points). In fact, the average affected dancer has gotten points at **`r mean_comps`** Novice competitions! Actually, now that I think about it, this makes sense -- most Novice dancers who get out via a small number of placements vs. one point at a time will overshoot 15.

Now let's break down of the number of 15-point-Novice-graduates in each region, taken as a percentage of total Novice grads. This is meant to give a measure of how much the rule change will impact each region going forward. 

```{r nov_byregion, echo=FALSE}
# Count total dancers per region who are not in Novice
# Use only those who graduated Novice under the 2015 rules
total_region <- new_novice %>% filter(division != 'Novice') %>% count(region)

# Count Novices with exactly 15 points in each region, take as % of total in each region
nov_byregion <- new_novice %>% 
  filter(nov_points == 15) %>% 
  count(region) %>%
  mutate(total = total_region$n) %>%
  mutate(percent = n/total*100) %>%
  mutate(SE = sqrt((n/total*(1-n/total))/total)*100)

# Plot data using ggplot2
ggplot(nov_byregion, aes(x=region, y=percent)) +
  # Error bars show standard error
  geom_errorbar(aes(ymin=percent-SE, ymax=percent+SE, width = 0.1)) +
  geom_bar(aes(fill=region), position='dodge', stat='identity') +
  xlab('Region') + ylab('% of total Novice-grads') +
  ggtitle('Proportion of Novice-grads with exactly 15 Novice points') +
  labs(fill="Region")

# Fisher test (is having 15 points related to region?)
# The Fisher test takes forever to run because we need such a large workspace (since our p-value and sample size are so low), but it's technically the correct one to use I think, so here it is.
test_data <- nov_byregion %>%
  mutate(yes = n) %>%
  mutate(no = total - n) %>%
  mutate(region = c(1,2,3,4,5)) %>%
  select(region, yes, no) %>%
  data.matrix()
p <- round(fisher.test(test_data, simulate.p.value=TRUE, B=1e7)$p.value, 3)
conf <- round(100*(1-p), 0)

# Mean number of Novice comps
new_novice %>%
  filter(nov_points == 15, int_points == 0) %>%
  group_by(region) %>%
  summarize(avg = round(mean(nov_count), 2))
```

It seems that the Midwest has a relatively lower proportion of dancers who escape Novice with 15 points exactly. When we Fisher test as we did in the previous section, we get a p-value of `r p` -- this suggests a `r conf`% confidence that **your likelihood to hit exactly 15 Novice points (and thereby be affected by the new Novice requirement) varies based on where you live.**

I'm kind of surprised, honestly. As for interpretation -- dancers in the Midwest go to fewer tier 3 Novice divisions maybe? A tier 3 Novice division gives out a single point more frequently than tier 2 (where only 5th-10th get a single point) and tier 1 (where only 5th place gets a single point), and is also the only way to get 15 points in one go. That's about all I can think of.

For good measure, I also broke down the mean number of pointed Novice events for 15-point-Novices by region. It's about the same across regions: 4-ish events. 


#### **How well do 15-point Novice dancers do in Intermediate?**
I decided to use two metrics for how "ready" a Novice dancer is for Intermediate. The first is **time to first Intermediate point**. It would be better to use number of competitions until first Intermediate point, but since the WSDC doesn't record points from competitions that dancers participate in but don't point at, I don't have access to this information. Additionally, I wanted to include dancers who don't yet have an Intermediate point (to catch dancers who have been competing in Intermediate but haven't gotten a point yet). To use that data, I pretended that these dancers will get their first point in January 2018.

The downside of using time rather then number of competitions is that it penalizes those who don't compete often. As a second metric robust to this bias, I looked at **ratio of placements to total** in Intermediate, with the logic that dancers who get most of their points from placing rather than making it to finals (but not placing) are better dancers compared to the rest of their division. Looking at just the number of placements wouldn't distinguish between placing at a tier 1 event and placing at a tier 3 event -- so, I chose to count entries with more than 5 points awarded -- this corresponds with 1st to 5th place at a tier 3 event or 1st to 3rd place at a tier 2 event. Of course, due to the nature of this dance/competition format, doing well in finals vs. prelims is not necessarily a measure of how good a dancer you are. Choosing good metrics is hard :(

I also broke down results by region, to account for any variation there.

```{r nov_readiness_time, echo=FALSE}
# Add variables for time to first intermediate point
# Use only those who graduated Novice under 2015 rules
nov_readiness_time <- new_novice %>%
  filter(nov_points >= 15) %>%
  mutate(has_15 = ifelse(nov_points == 15, 1, 0)) %>% # Indicator variable = 1 if 15 Novice points
  mutate(time_to_int = ifelse(int_points > 0, 
                              as.integer(as.Date(paste(int_start, '01'), '%B %Y %d')  - as.Date(paste(nov_end, '01'), '%B %Y %d'))/30, 
                              as.integer(as.Date('January 2018 01', '%B %Y %d') - as.Date(paste(nov_end, '01'), '%B %Y %d'))/30)) %>%
# I noticed some negative values of time_to_int -- these could be caused by combined nov/int competitions being recorded as Novice? Or just some other error in the database. I'm gonna just filter these out.
  filter(time_to_int >= 0)

# Summarize data for plotting
# It's pretty skewed so I'm using medians
plot_data <- nov_readiness_time %>%
  group_by(has_15, region) %>%
  summarize(median = median(time_to_int))

# Build plot for time to first intermediate point by region and if(single novice comp)
ggplot(plot_data, aes(x=region, y=median, fill=as.factor(has_15))) +
  geom_bar(position=position_dodge(0.9), stat='identity') +
   xlab('Region') + ylab('Median months until first Intermediate point') +
  scale_fill_discrete(name='Exactly 15 Novice points?',
                      breaks=c('0', '1'),
                      labels=c('No', 'Yes'))

# Medians and Mann-Whitney tests across regions
medians <- nov_readiness_time %>% group_by(has_15) %>% summarize(median=median(time_to_int))
to_print <- c(medians$median[2], medians$median[1], wilcox.test(time_to_int~has_15, data=nov_readiness_time)$p.value)
names(to_print) <- c('15 Points', '16+ Points', 'p-value')
to_print
```

Because the medians of each region are comparable to each other, I decided to look at combined data from all the regions. Across regions, we find >99% confidence that dancers with exactly 15 Novice points **take longer on average to get their first Intermediate point**. 

On the other hand, the difference in medians is only about 2 months, so how much does that really matter?

[Stats note: quick histograms of the data (not shown) revealed it's pretty skewed across all regions, hence I've used medians and nonparametric hypothesis testing]

```{r nov_readiness_ratio, echo=FALSE}
# Add variables for ratio of placements to total
# Use only those who graduated Novice under 2015 rules
nov_readiness_ratio <- new_novice %>%
  filter(int_count > 0) %>%
  mutate(has_15 = ifelse(nov_points == 15, 1, 0)) %>% # Indicator variable = 1 if only 1 novice comp before intermediate
  mutate(place_ratio = int_place/int_count)

# Summarize data for plotting
# Again, these data are pretty skewed, so we'll use medians
plot_ratio <- nov_readiness_ratio %>%
  group_by(has_15, region) %>%
  summarize(median = median(place_ratio))

# Build plot for ratio of placements to total
ggplot(plot_ratio, aes(x=region, y=median, fill=as.factor(has_15))) +
  geom_bar(position=position_dodge(0.9), stat='identity') +
   xlab('Region') + ylab('Median ratio of placements to total') +
  scale_fill_discrete(name='Exactly 15 Novice points?',
                      breaks=c('0', '1'),
                      labels=c('No', 'Yes'))

# Mann-Whitney tests by region
lapply(split(nov_readiness_ratio, nov_readiness_ratio$region), function(x)with(x, wilcox.test(place_ratio~has_15)$p.value))

```

Well, it looks like the low incidence of placing 1st-5th at a tier 3 or 1st-3rd at a tier 2 has maybe come back to bite us -- this data is decidedly not as pretty.

Significance testing by region, we see a 93% confidence that dancers in the Other region (Asia/Australia/Brazil) with exactly 15 Novice points **place less frequently in Intermediate**. The is also the trend in 3 of the other 4 regions (East Coast, Europe, Midwest), although the differences there are less significant. 

Combining the two metrics, there is some evidence that Novice dancers with exactly 15 points do slightly worse in Intermediate than those with more than 15 points. Although it's not a large difference, and changing the requirement from 15 to 16 points is unlikely to do much, this does justify the rule change somewhat.


### **Community growth in the US and overseas**

One question that's come up in the discussions surrounding the rule changes is whether the decisions were made in a US-centric way by a board that contains only members who live in the US/don't travel enough. The argument has been made that West Coast Swing has grown into an international dance, and that the WSDC board should have more international representation. I tried to use WSDC data to see just how international the dance has become/is becoming.


#### **Percentage of active competitors by region**

The most direct way to look at growth is to see what percentage of active competitors (at least 1 point in the last 2 years) live in each region. I'm about 90% sure that I did this in the stupidest possible way, but hey, the plots work so I think that deserves partial credit.

```{r percent_byregion, fig.width=8, echo=FALSE}
# Tracking percent of total actively competing dancers who live in each region
percent_byregion <- data %>%
  mutate(all_start = as.Date(paste(ifelse(is.na(nov_start), new_start, nov_start), '01'), '%B %Y %d')) %>%
  group_by(region) %>%
  count(all_start) %>%
  mutate(count = cumsum(n)) %>%
  filter(all_start >= as.Date('2012-01-01'))

# Get total number of dancers across regions by date
totals <- data %>%
  mutate(all_start = as.Date(paste(ifelse(is.na(nov_start), new_start, nov_start), '01'), '%B %Y %d')) %>%
  count(all_start) %>%
  mutate(total_count = cumsum(n)) %>%
  filter(all_start >= as.Date('2012-01-01'))

# Join the data frames together to calculate percent of total for each region
totals <- bind_rows(totals, totals, totals, totals, totals)
totals <- totals %>% mutate(region = c(rep('East Coast', 72), rep('Europe', 72), rep('Midwest', 72), rep('Other', 72), rep('West Coast', 72)))
percent_byregion <- left_join(totals, percent_byregion, by = c("all_start", "region"))
filled <- percent_byregion %>%
  fill(count)
percent_byregion <- filled %>%
  mutate(percent = count / total_count * 100) %>%
  select(all_start, region, percent)

# Combine all the US regions
US_percent <- percent_byregion %>%
  group_by(all_start) %>%
  filter(region %in% c('West Coast', 'East Coast', 'Midwest')) %>%
  summarize(percent = sum(percent)) %>%
  mutate(region = 'US (total)')

# Bind with Europe and Other regions
percent_byregion2 <- bind_rows(US_percent, filter(percent_byregion, region %in% c('Europe', 'Other')))

# Build plots
all_regions <- ggplot(percent_byregion, aes(x=all_start, y=percent, fill=region)) +
  geom_bar(position='stack', stat='identity') +
  #coord_cartesian(xlim = as.Date(c('2013-01-01', '2017-12-01'))) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
  labs(fill = 'Region', title = 'Percent of active competitors in each region')

combined_US <- ggplot(percent_byregion2, aes(x=all_start, y=percent, fill=region)) +
  geom_bar(position='stack', stat='identity') +
  coord_cartesian(xlim = as.Date(c('2013-01-01', '2017-12-01'))) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
  labs(fill = 'Region')

plot_grid(all_regions, combined_US, align = 'v', ncol = 1, labels = NULL, axis = 'b')
```

The share of competitors living in each of the 3 regions of the US (East Coast, Midwest, West Coast) has shrunk slightly since 2013, while the proportions living in Europe and Asia/Australia/Brazil (Other) have both increased.

Bundling all of the US regions into one, it becomes clearer that international competitors have been quite a large portion of the total for some time, and that this fraction continues to grow. International competitors have comprised **over 30%** of the total since mid-2014 (over 3 years ago!), and currently represent **38%** of all competitors. By this measure, international representation on the WSDC board certainly seems overdue.


#### **Growth rate of number of dancers by region**

In trying to write this, I discovered that it's more difficult than I thought to estimate the absolute number of dancers in a region from the points registry data (which can only give you the number of pointed competitors). The main problem is that the number of pointed competitors in a region depends both on how many dancers there are *and* how many times points were awarded in that region(i.e., how many events and what tier the events are). 

I decided to total the number of times a point was awarded by an event in each region in 2014-2018 (using eventregions.py) and divide the number of competitors by that quantity. This is used as a proxy that depends much less on number of events and size of events in the region.

```{r count_byregion, fig.width=8, echo=FALSE}
# Tracking total number of active dancers per region
count_byregion <- data %>%
  mutate(all_start = as.Date(paste(ifelse(is.na(nov_start), new_start, nov_start), '01'), '%B %Y %d')) %>%
  group_by(region) %>%
  count(all_start) %>%
  mutate(total_count = cumsum(n))

# Make cumulative US data
US_data <- data %>%
  mutate(all_start = as.Date(paste(ifelse(is.na(nov_start), new_start, nov_start), '01'), '%B %Y %d')) %>%
  mutate(region = ifelse(region %in% c('Europe', 'Other'), region, 'US (net)')) %>%
  group_by(region) %>%
  count(all_start) %>%
  mutate(total_count = cumsum(n))

# Add to the rest of the data
count_byregion <- bind_rows(count_byregion, US_data)

# Tibble of point-award-instance count by region in 2014 and later
event_count <- tibble(region = c('Other', 'Europe', 'Midwest', 'West Coast', 'US (net)', 'East Coast'), count = c(1160, 5415, 3078, 3878, 10294, 3338))

# Join to the rest of the data
count_byregion <- left_join(count_byregion, event_count, by='region')
count_byregion <- count_byregion %>%
  # Dividing by number of point-award-instances in region
  mutate(total_count = (total_count / count)) %>%
  mutate(months = as.integer(all_start - as.Date('2014-01-01'))/30) %>%
  filter(months > 0) %>%
  ungroup() %>%
  mutate(region = as.factor(region))

# Re-order regions
count_byregion$region <- factor(count_byregion$region, levels(count_byregion$region)[c(4, 2, 3, 5, 6, 1)])

# Build plot
ggplot(count_byregion, aes(x=months, y=total_count, color=region)) +
  geom_line(linetype=2) +
# Regressions; growth is relatively linear in the last 4 years, so we'll just fit a linear model to the last 4 years.
  geom_smooth(method = 'lm') +
  stat_poly_eq(formula = y ~ x,
               aes(label =  ..eq.label..), 
               parse = TRUE) +
  scale_color_discrete(name='Region') +
  xlab('Months since January 2014')+ ylab('Active competitors per in-region record') + 
  ggtitle('Growth by region')

```

(Dotted line shows actual data; solid line shows linear fit. By "in-region record", I mean "instance where an event in that region awarded points to someone in 2014-2018")

I think the closest I can come to intuiting the value being plotted is to think about it as a ratio of the number of competitors to the number of events (weighted by event size, since tier 3 events give points to more people than tier 2 events and so on) in a region. A higher value indicates 1) the dancers in that region earn points outside of their own region more frequently and 2) the events in that region award points to more new dancers (as opposed to dancers who already have a WSDC number).

The growth rate of this value should be a pretty good descriptor of the growth rate of each dance community, so here they are from highest to lowest:

* Other: 0.00552 [competitors/record/month, not that this unit means much]
* Europe: 0.00434 
* Midwest: 0.00404
* US (net): 0.00375
* West Coast: 0.00372
* East Coast: 0.00352

The estimated growth rate in Asia/Australia/Brazil is about 47% higher than that of the US, and Europe is growing about 16% faster than the US. The takeaway is that international representation on WCS decision-making boards will be even more important as time goes on.


## **Dataset**
Data was scraped from the WSDC website and processed by wsdcscraper.py. The Python code generates a .csv file, which was then read into R for the rest of the analysis. Only data from dancers with at least 1 Novice point were collected.


### **Variables**

* id: WSDC ID number
* role: Primary role (leader/follower)
* region: Estimated region of residence
* end_date: Date of most recent point entered
* nov_points: Number of points in Novice
* nov_start: Date of first Novice point
* nov_end: Date of last Novice point
* nov_count: Number of events with Novice points recorded
* nov_place: Number of (1-5) placements in Novice
* nov_first: Indicates whether placed first in Novice

The last 6 variables are duplicated for Intermediate (int), Advanced (adv), and All-Star (als).


### **Regions**
Since regional differences were a common theme in the discussions, I tried to estimate what region each dancer lives in by looking at where the events that they pointed at are located (the logic is that people will mostly travel to and get points from events close to them). Here are the region definitions:

* West Coast: West Coast US plus British Columbia
* Midwest+TX: Midwest US
* East Coast: East Coast US plus the rest of Canada
* Other: Singapore, South Korea, Australia, New Zealand, Brazil
* Europe: All else (...including Russia and Israel*)

For dancers in the US and Canada, I used whichever region they had the most events in. Since dancers from Europe or isolated communities often travel to other regions to compete, I defined those differently. Any dancer with more than 5 European events or at least 1 European event and less than 10 events total was identified as a European dancer. Any dancer with more than 3 events in an isolated region or at least 1 such event and less than 10 events total was identified as the "other" region. These estimations are, of course, not perfect, but a quick spot check of dancers I know deem them to be good enough for me! It looks like the largest bias is towards being identified as a "West Coast" dancer if you live in the Midwest/TX, since there are relatively fewer events in that region and it's a relatively short flight to the West Coast, where there are more/larger events. 

*One thing I'm not sure about -- should Russia and Israel count as Europe? I wanted to separate out Brazil, Asia, and Australia because of discussions after the recent rule change about how these communities are more isolated, as it's more expensive to travel from there to the rest of the world and the local WSDC scene is not yet robust. I feel like Russia and Israel are not quite so isolated but I'm not 100% sure -- anyone have input on this?